Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-02-04 src/tex/ms.tex
added some discussion in 4.1 addressing Reviewers 1 comment number 2

\section{Introduction}
\label{sec:intro}
A wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
This “UV Threshold Hypothesis” states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.

@@

@@

\rev{
\subsubsection{Other mechanisms regulating habitability and abiogenesis}
Our study focuses primarily on two factors that may regulate the emergence and persistence of life on a planet: the \gls{NUV} flux and the viability of liquid water, which provide a testable framework for assessing biosignature distributions.
Of course, planetary habitability and abiogenesis depend on a broader range of physical and chemical conditions.
For one, a planet must retain an atmosphere capable of supporting liquid water and surface chemistry.
Atmospheric escape processes have been studied in detail, and their occurrence is supported by planet formation models~\citep[e.g.,][]{Owen2013,Schlichting2014a,Ginzburg2016b,Mordasini2020,Burn2024} as well as exoplanet demographics~\citep[e.g.,][]{Owen2019a,Bergsten2022,Rogers2021a}.
Population synthesis studies demonstrate that models of atmospheric escape can explain key statistical features of the observed exoplanet population~\citep{Rogers2020,Emsenhuber2021b,Schlecker2021b,Burn2024}.}

\rev{
Planets around M~dwarfs, in particular, may be subject to significant atmospheric loss due to the extended high-luminosity pre-main-sequence phase of their host stars~\citep[e.g.,][]{Luger2015}.
%During this period, extreme X-ray and UV irradiation can drive rapid hydrodynamic escape, potentially stripping a planet of its primordial atmosphere before abiogenesis can occur.
%Even after this early phase, M-dwarf planets remain exposed to intense stellar activity, leading to further erosion of secondary atmospheres if not replenished through volcanic outgassing or other mechanisms.
Recent works~\citep[e.g.,][]{ParkCoy2024,Luque2024} interpret JWST eclipse measurements as growing evidence for the absence of substantial atmospheres on M-dwarf rocky exoplanets.
If these planets indeed lack atmospheres, their habitability would be primarily dictated by the rate of atmospheric escape~\citep[e.g.,][]{Owen2020}.
However, alternative interpretations remain viable: \citet{Ducrot2024} and \citet{Hammond2025} recently demonstrated that current JWST data cannot definitively distinguish between a bare-rock scenario and an atmosphere composed primarily of N$_2$-CO$_2$-H$_2$O.
This underscores the need for future observations and modeling efforts to constrain atmospheric retention and escape processes more robustly.}
...

%For instance, atmospheric escape processes may play a critical role not only in maintaining a planet's atmosphere but also in regulating the planet's water inventory and the availability of volatile elements for prebiotic chemistry.

\rev{While not modeled here, the above additional mechanisms may offer directions for future work.}

>>>>

\section{Introduction}
\label{sec:intro}
\rev{While the probability of abiogenesis is unknown and has primarily been explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024},
a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
This “UV Threshold Hypothesis” states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.

@@

\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
\begin{equation}
    \label{eq:BayesTheorem}
    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
\end{equation}
However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
Taking the ratio of the posterior evidence for two alternate hypotheses (i.e. of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
\begin{equation}
    \label{eq:postratio}
    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
\end{equation}
This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}

\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
\begin{equation}
    \label{eq:BayesTheorem}
    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
\end{equation}
However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
Taking the ratio of the posterior evidence for two alternate hypotheses (i.e. of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
\begin{equation}
    \label{eq:postratio}
    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
\end{equation}
This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}

@@

\rev{
\subsubsection{Other mechanisms regulating habitability and abiogenesis}
Our study focuses primarily on two factors that may regulate the emergence and persistence of life on a planet: the \gls{NUV} flux and the viability of liquid water, which provide a testable framework for assessing biosignature distributions.
Of course, planetary habitability and abiogenesis depend on a broader range of physical and chemical conditions.
For one, a planet must retain an atmosphere capable of supporting liquid water and surface chemistry.
Atmospheric escape processes have been studied in detail, and their occurrence is supported by planet formation models~\citep[e.g.,][]{Owen2013,Schlichting2014,Ginzburg2016,Mordasini2020,Burn2024} as well as exoplanet demographics~\citep[e.g.,][]{Owen2019,Bergsten2022,Rogers2021}.
Population synthesis studies demonstrate that models of atmospheric escape can explain key statistical features of the observed exoplanet population~\citep{Rogers2020,Emsenhuber2021b,Schlecker2021b,Burn2024}.}

\rev{
Planets around M~dwarfs, in particular, may be subject to significant atmospheric loss due to the extended high-luminosity pre-main-sequence phase of their host stars~\citep[e.g.,][]{Luger2015}.
%During this period, extreme X-ray and UV irradiation can drive rapid hydrodynamic escape, potentially stripping a planet of its primordial atmosphere before abiogenesis can occur.
%Even after this early phase, M-dwarf planets remain exposed to intense stellar activity, leading to further erosion of secondary atmospheres if not replenished through volcanic outgassing or other mechanisms.
Recent works~\citep[e.g.,][]{Coy2024,Luque2024} interpret JWST eclipse measurements as growing evidence for the absence of substantial atmospheres on M-dwarf rocky exoplanets.
If these planets indeed lack atmospheres, their habitability would be primarily dictated by the rate of atmospheric escape~\citep[e.g.,][]{Owen2020}.
However, alternative interpretations remain viable: \citet{Ducrot2024} and \citet{Hammond2025} recently demonstrated that current JWST data cannot definitively distinguish between a bare-rock scenario and an atmosphere composed primarily of N$_2$-CO$_2$-H$_2$O.
This underscores the need for future observations and modeling efforts to constrain atmospheric retention and escape processes more robustly.}
%For instance, atmospheric escape processes may play a critical role not only in maintaining a planet's atmosphere but also in regulating the planet's water inventory and the availability of volatile elements for prebiotic chemistry.

\rev{Beyond atmospheric escape, internal heating from tidal forces or radioactive decay can extend or constrain the limits of planetary surface habitability~\citep[e.g.,][]{Barnes2013,Oosterloo2021}.
Tidal effects are especially relevant for \gls{HZ} M~dwarf planets, where strong stellar interactions can drive internal heating, loss of water or an atmosphere, or runaway greenhouse conditions.
Tidal locking may also create extreme climate zones, challenging habitability.}
\rev{While not modeled here, the factors outlined above may offer directions for future work.}
